---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# intFRT

<!-- badges: start -->
<!-- badges: end -->

The goal of **intFRT** is to integrate randomized controlled trials (RCTs) with external controls (ECs) in hybrid controlled trials, harnessing Fisher randomization tests (FRT) and Conformal Selective Borrowing (CSB). It enhances the statistical efficiency of average treatment effect (ATE) estimation and inference while ensuring valid hypothesis testing. Key features include:

- Fisher randomization tests (FRT) in hybrid controlled trials, ensuring strict control of the Type I error rate.

- Conformal Selective Borrowing (CSB) serves as both an ATE estimator and a test statistic for FRT, enabling selective borrowing of comparable ECs to reduce hidden bias and improve statistical power.

- Adaptive determination of the selection threshold for CSB.

- The package also provides No Borrowing and various EC borrowing estimators (IPW, staIPW, CW, OM, AIPW, ACW), along with their inference results based on asymptotic normality for comparison.

## Installation

You can install the development version of intFRT from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("ke-zhu/intFRT")
```

## Example

This example illustrates the use of the Fisher Randomization Test (FRT) and Conformal Selective Borrowing methods in hybrid controlled trials.

### Simulate data for a hybrid controlled trial

```{r}
set.seed(1)
n_rct <- 50  # Number of observations in the randomized controlled trial
n_ec <- 100  # Number of external controls
n <- n_rct + n_ec  # Total number of observations

# Covariates (2 covariates, uniformly distributed)
X <- matrix(runif(n * 2), n, 2)

# Treatment assignment (1 = treatment, 0 = control)
A <- c(rep(0:1, each = n_rct / 2), rep(0, n_ec))

# Data source indicator (1 = randomized trial, 0 = external control)
S <- c(rep(1, n_rct), rep(0, n_ec))

# Generate potential outcomes (continuous)
Y1 <- 1 + 2 * X[,1] + 0.1 * X[,2] + rnorm(n)
Y0 <- 2 * X[,1] + 0.1 * X[,2] + rnorm(n)

# Introduce bias in half of the external controls
id_biased_EC <- tail(which(S == 0), n = n_ec / 2)
Y0[id_biased_EC] <- Y0[id_biased_EC] - 10

# Observed outcome
Y <- A * Y1 + (1 - A) * Y0
```


### Adaptive selection threshold

```{r}
library(intFRT)
ada_g <- compute_ada_gamma(
  Y, A, S, X, 
  # Use a small n_rep_gamma for fast illustration; 
  # recommend n_rep_gamma = 100 for more stable results
  n_rep_gamma = 10
)
ada_g
```


### Fisher Randomization Test with Conformal Selective Borrowing

```{r}
result_csb <- ec_borrow(
  Y = Y,
  A = A,
  S = S,
  X = X,
  method = "Conformal Selective Borrow AIPW",
  family = "gaussian",
  gamma_sel = ada_g,
  # FRT with 10 randomizations for illustration
  # recommend n_fisher = 5000 for more powerful results
  n_fisher = 10
)

# View results
print(result_csb$res)

# View IDs of borrowed external controls
result_csb$out$id_sel[[1]]
```


### Visualize borrowed external controls

```{r example, fig.width=7, fig.height=5, dev='svg'}
library(dplyr)
library(ggplot2)
library(forcats)
library(quantreg)

# Mark borrowed external controls for plotting
sel <- rep(0, length(Y))
sel[result_csb$out$id_sel[[1]]] <- 1

# Fit sampling score for multiple covariates 
`Sampling Score` <- glm(S ~ X) %>% predict(type = "response")

# Prepare data for plotting
dat_plot <- tibble(Y, A, S, X, `Sampling Score`, sel) %>% 
  mutate(
    Type = case_when(
      A == 1 & S == 1 ~ "RCT treated",
      A == 0 & S == 1 ~ "RCT control",
      A == 0 & S == 0 & sel == 1 ~ "External control (selected)",
      A == 0 & S == 0 & sel == 0 ~ "External control (unselected)"
    ) %>% 
      as_factor %>% 
      fct_relevel(
        "RCT treated", "RCT control", 
        "External control (selected)", "External control (unselected)"
      )
  ) %>% 
  filter(Type != "RCT treated")

# Fit quantile regressions to visualize the main range of RCT controls
fit975 <- rq(Y ~ `Sampling Score`, tau = 0.975, data = dat_plot, 
             subset = dat_plot$Type == "RCT control")
dat_plot$pred975 <- predict(fit975, newdata = dat_plot)

fit025 <- rq(Y ~ `Sampling Score`, tau = 0.025, data = dat_plot, 
             subset = dat_plot$Type == "RCT control")
dat_plot$pred025 <- predict(fit025, newdata = dat_plot)

# Plot results
dat_plot %>% 
  ggplot() +
  geom_ribbon(aes(`Sampling Score`, ymin = pred025, ymax = pred975), 
              fill = "grey80", alpha = 0.5) +
  geom_point(aes(`Sampling Score`, Y, color = Type)) +
  scale_color_manual(values = c("#5A5A5A", "#00ADFA", "#F8766D"))
```


