---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# intFRT

<!-- badges: start -->
<!-- badges: end -->

The goal of **intFRT** is to **int**egrate randomized controlled trials (RCTs) with external controls (ECs) in hybrid controlled trials, harnessing Fisher randomization tests (**FRT**) and Conformal Selective Borrowing (CSB). It enhances the statistical efficiency of average treatment effect (ATE) estimation and inference while ensuring valid hypothesis testing. Key features include:

🎲 **Fisher Randomization Tests (FRT)**: Ensures strict control of the Type I error rate in hybrid controlled trials.

🧲 **Conformal Selective Borrowing (CSB)**: Serves as both an ATE estimator and a test statistic for FRT, selectively borrowing comparable ECs to reduce hidden bias and enhance statistical power.

⚙️ **Adaptive Selection Threshold**: Dynamically determines the selection threshold for CSB.

📊 **Flexible Estimator Options**: Offers No Borrowing and various EC borrowing estimators (IPW, staIPW, CW, OM, AIPW, ACW) with inference results based on asymptotic normality. These estimators can also be used as test statistics in FRT.



## Installation

You can install the development version of intFRT from [GitHub](https://github.com/) with:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("ke-zhu/intFRT")
```

## Example

This example illustrates the use of the Fisher Randomization Test (FRT) and Conformal Selective Borrowing methods in hybrid controlled trials.

### Simulate data for a hybrid controlled trial

```{r sim}
set.seed(2025)
n_rct <- 50  # Number of observations in the randomized controlled trial
n_ec <- 100  # Number of external controls
n <- n_rct + n_ec  # Total number of observations

# Covariates (2 covariates, uniformly distributed)
X <- matrix(runif(n * 2), n, 2)

# Treatment assignment (1 = treatment, 0 = control)
A <- c(rep(0:1, each = n_rct / 2), rep(0, n_ec))

# Data source indicator (1 = randomized trial, 0 = external control)
S <- c(rep(1, n_rct), rep(0, n_ec))

# Generate potential outcomes (continuous)
Y1 <- 1 + 2 * X[,1] + 0.1 * X[,2] + rnorm(n)
Y0 <- 2 * X[,1] + 0.1 * X[,2] + rnorm(n)

# Introduce bias in half of the external controls
id_biased_EC <- tail(which(S == 0), n = n_ec / 2)
Y0[id_biased_EC] <- Y0[id_biased_EC] - 10

# Observed outcome
Y <- A * Y1 + (1 - A) * Y0
```


### Adaptive selection threshold

```{r ada_g}
library(intFRT)
ada_g <- compute_ada_gamma(
  Y, A, S, X, 
  # Tuning with 20 replications for illustration purposes
  # Recommend setting `n_rep_gamma = 100` or higher with parallel computing
  n_rep_gamma = 20
)
ada_g
```


### Fisher Randomization Test with Conformal Selective Borrowing

```{r frt}
result_csb <- ec_borrow(
  Y = Y,
  A = A,
  S = S,
  X = X,
  method = "Conformal Selective Borrow AIPW",
  family = "gaussian",
  gamma_sel = ada_g,
  # FRT with 10 randomizations for illustration
  # Recommend `n_fisher = 5000` or more
  # To perform only Conformal Selective Borrowing, set `n_fisher = NULL`
  n_fisher = 10
)

# View results
print(result_csb$res, width = Inf)
```

- `est`: ATE estimate.
- `se`, `ci_l`, `ci_u`, `p_value` (first row): Standard error, confidence interval, p-value based on asymptotic normality.
- `p_value` (second row): P-value based on FRT.
- `n_sel`: The number of borrowed external controls.
- `ess_sel`: The effective sample size of borrowed external controls.
- `runtime`: The computation time.

```{r ecs}
# View IDs of borrowed external controls
result_csb$id_sel
```


### Visualize borrowed external controls

```{r vis, fig.width=7, fig.height=5, dev='svg'}
library(dplyr)
library(ggplot2)
library(forcats)
library(quantreg)

# Mark borrowed external controls for plotting
sel <- rep(0, length(Y))
sel[result_csb$id_sel] <- 1

# Fit sampling score for multiple covariates 
`Sampling Score` <- glm(S ~ X) %>% predict(type = "response")

# Prepare data for plotting
dat_plot <- tibble(Y, A, S, X, `Sampling Score`, sel) %>% 
  mutate(
    Type = case_when(
      A == 1 & S == 1 ~ "RCT treated",
      A == 0 & S == 1 ~ "RCT control",
      A == 0 & S == 0 & sel == 1 ~ "External control (selected)",
      A == 0 & S == 0 & sel == 0 ~ "External control (unselected)"
    ) %>% 
      as_factor %>% 
      fct_relevel(
        "RCT treated", "RCT control", 
        "External control (selected)", "External control (unselected)"
      )
  ) %>% 
  filter(Type != "RCT treated")

# Fit quantile regressions to visualize the main range of RCT controls
fit975 <- rq(Y ~ `Sampling Score`, tau = 0.975, data = dat_plot, 
             subset = dat_plot$Type == "RCT control")
dat_plot$pred975 <- predict(fit975, newdata = dat_plot)

fit025 <- rq(Y ~ `Sampling Score`, tau = 0.025, data = dat_plot, 
             subset = dat_plot$Type == "RCT control")
dat_plot$pred025 <- predict(fit025, newdata = dat_plot)


# Plot results
dat_plot %>% 
  ggplot() +
  geom_ribbon(aes(`Sampling Score`, ymin = pred025, ymax = pred975), 
              fill = "grey80", alpha = 0.5) +
  geom_point(aes(`Sampling Score`, Y, color = Type)) +
  scale_color_manual(values = c("#5A5A5A", "#00ADFA", "#F8766D"))
```

The above figure shows that Conformal Selective Borrowing discards incomparable ECs and borrows most of the comparable ones, except for a few suspicious ones.


## Reference

Ke Zhu, Shu Yang, and Xiaofei Wang (2025). [Enhancing statistical validity and power in hybrid controlled trials: A randomization inference approach with conformal selective borrowing.](https://arxiv.org/abs/2410.11713) *Proceedings of the 42nd International Conference on Machine Learning (ICML)*, PMLR, in press. [[Slides]](https://drive.google.com/file/d/1LkTDY12CjUL0BAGQAt4JmOBlgk-MN3Ix/view?usp=sharing) [[Poster]](https://drive.google.com/file/d/1g5vFT6irtPWFQWwh6AGe-iYCvMF4z0B2/view?usp=share_link)
